---
title: "lab4"
author: "HsiHsuan Yang"
format:
  html:
    embed-resources: true
---
```{r}
library(R.utils)
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```
Question 2
```{r}
library(data.table)
library(leaflet)
library(tidyverse)

# Step 1: Remove temperatures less than -17C
met <- met[temp >= -17]

# Step 2: Replace missing data (coded as 9999, 999, etc.) with NA
met[elev == 9999, elev:= NA]

# Step 3: Generate a date variable using the given hint
met[, week := week(as.Date(paste(year, month, day, sep = "-")))]

# Step 4: Keep observations of the first week of the month using week function
met <- met[week == min(week, na.rm = TRUE)]

#Step 5: average of each station
met_mean <- met[,.(temp=mean(temp,na.rm=TRUE), rh=mean(rh,na.rm=TRUE), windsp=mean(wind.sp,na.rm=TRUE), 
                visdist=mean(vis.dist,na.rm=TRUE), dewpoint = mean(dew.point, na.rm=TRUE), lat=mean(lat), lon=mean(lon), 
                elev=mean(elev,na.rm=TRUE)), by="USAFID"]

met_mean$elev_cat <- ifelse(met_mean$elev> 252, "high", "low")

met_mean$region <- ifelse(met_mean$lon > -98 & met_mean$lat >39.71, "NE",
                         ifelse(met_mean$lon > -98 & met_mean$lat < 39.71, "SE",
                                ifelse(met_mean$lon < -98 & met_mean$lat >39.71, "NW", "SW")))

table(met_mean$region)
```
Question 3
```{r}
met_mean %>%
  filter(!(region %in% NA)) %>%
ggplot()+
  geom_violin(mapping = aes(y=windsp, x=1)) +
  facet_wrap(~region, nrow=2)
```
Question 4
```{r}
met_mean %>%
filter(!(region %in% NA)) %>%
  ggplot(mapping = aes(x=dewpoint, y=rh, color=region))+
  geom_jitter() + 
  stat_smooth(method=lm)
```
Question 5
```{r}
met_mean %>%
filter(!(region %in% NA)) %>%
  ggplot()+
  geom_bar(mapping=aes(x=elev_cat,fill=region), position = "dodge")+
  scale_fill_brewer(palette = "Pastel1")+
  labs( x="elevation category", y= "count")+
  theme_bw()
```
Question 6
```{r}
met_mean %>%
filter(!(region %in% NA)) %>%
  ggplot(mapping=aes(x=region, y=dewpoint)) +
  stat_summary(fun.data="mean_sdl", geom="errorbar") +
  stat_summary(fun.data="mean_sdl")
met_mean %>%
filter(!(region %in% NA)) %>%
  ggplot(mapping=aes(x=region, y=windsp)) +
  stat_summary(fun.data="mean_sdl", geom="errorbar") +
  stat_summary(fun.data="mean_sdl")
```
Based on the above result, the mean dew point for NE is around 16, NW is around 12.5, SE is around 21, SW is around 14. The wind speed for NE is around 1.75, NW is around 2.8, SE is around 1.75, SW is around 3.1.  

Question 7
```{r}
met_mean2<-met_mean[!is.na(rh)]

# Top five
top5 <- met_mean2[rank(-rh) <= 10]

rhpal = colorNumeric(c('pink','blue','green'), domain=met_mean2$rh)
leaflet(met_mean2) %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat=~lat, lng=~lon, color=~rhpal(rh), label=~paste0(round(rh,2), ' rh'), opacity=1,fillOpacity=1, radius=500) %>%
  addMarkers(lat=~lat, lng=~lon, label=~paste0(round(rh,2), ' rh'), data = top5) %>%
  addLegend('bottomleft',pal=rhpal, values=met_mean2$rh, title="RH", opacity=1)
```
The RH is the highest on the east coast and lowest on the west coast (besides a few places near the coast of the westcoast that have high RH) of the United States according to the map.

Question 8-Attempt 1
```{r}
library(ggforce)
ggplot(met_mean, aes(elev, temp, colour = region)) +
  geom_point() +
  facet_zoom(x = region == "versicolor")
```








